---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gapr

<!-- badges: start -->
<!-- badges: end -->

gapr is, at the moment, a small package designed to estimate distributional gaps in terms of an effect sizes, along with the associated standard error, using only count data from each distribution. This assumes that the counts are from ordered bins. The ECDF from each distribution is first approximated using these counts (cumulative proportion within each bin). The ECDF's are then paired and a smoothed curve is estimated using maximum likelihood. The area under the curve is then estimated and transformed to an effect size. The variance-covariance matrix associated with the curve fit is also used to estimate the standard error of the effect size. For more information, please see [Ho & Reardon (2012)](https://journals.sagepub.com/doi/abs/10.3102/1076998611411918?journalCode=jebb) and [Reardon & Ho (2015)](https://journals.sagepub.com/doi/10.3102/1076998615570944).

## Installation
gapr is currently only available on GitHub

``` r
# If not previously installed, first install the remotes package
# install.packages("remotes")

remotes::install_github("datalorax/gapr")
```

## Example

Estimate an achievement gap for a given school as follows

```{r example}
library(gapr)
ashland_middle_g6 <- oregon_schools[1:4, ]
ashland_middle_g6
```
The above represents the number of students scoring in each of 4 ordered 
proficiency categories on the Oregon statewide test of English/Language Arts 
in 2014-15 in Ashland Middle School for Grade 6, reported separately by
whether the student was coded as Hispanic/Latino or White.

We can estimate the difference between these student groups using the 
`estimate_v` function, and the resulting estimate (representing an effect size)
effectively recovers the distributional differences *as if we had the full
student-level data*.

```{r estimate-v}
estimate_v(ashland_middle_g6, "white", "hispanic_latino")
```

```{r estimate-v-store, echo = FALSE}
v <- estimate_v(ashland_middle_g6, "white", "hispanic_latino")
```

So, in this grade, at this school, students coded Hispanic/Latino scored, on
average, approximately `r round(abs(v$v), 2)` standard deviations below students coded
White.

Currently, the function only estimates for one school at a time, but future
developments will include `by` arguments to estimate gaps by another variable
(e.g., schools, districts).

## Code of Conduct

Please note that the gapr project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
